Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-01-29 parallel-tractability-J-v9/parallel-tractability-J.tex

\address[add2]{
    Institution of Artificial Intelligence, University of Ulm
}

\begin{abstract}
Materialization is an important reasoning service for applications built
on the ontology languages, such as OWL. With a
rapid growth of semantic data, it is challenging to perform
materialization on large-scale ontologies efficiently.
Current work proposes parallel algorithms and employs computing
platforms of high-performance to make materialization sufficiently efficient and scalable in practice.
However, there do exist ontologies, for which
materialization cannot always be improved by parallelism (we use experiments to show this in Section~\ref{sec:evaluation}).
On the other hand, some well-known large-scale ontologies, such as YAGO, have been shown to have
good performance for parallel reasoning, but they are expressed in ontology languages
that are not parallelly tractable in theory, i.e., the efficiency of reasoning may not be improved
on a parallel implementation.
This motivates us to study the problem of \emph{parallel tractability} of ontology
materialization from a theoretical perspective.
That is, we aim to identify the ontologies for which materialization is
parallelly tractable, i.e., in NC complexity.
The results of the parallel tractability can be further used to optimize
materialization algorithms and to guide ontology engineers in creating parallelly tractable ontologies.

In this work, we focus on datalog rewritable ontology languages.
We propose some algorithms, called NC algorithms, to identify several classes of datalog rewritable
ontologies (called \emph{parallelly tractable classes})
such that materialization over them is parallelly tractable.
We then investigate the parallel tractability of materialization over
two datalog rewritable ontology languages DL-Lite and DHL (\emph{Description Horn Logic}).
We show that materialization over any ontology expressed in DL-Lite$_{core}$ or DL-Lite$_\mathcal{R}$ is parallelly tractable.
For DHL, there exist ontologies where materialization can hardly be parallelized.
We propose to restrict the usage of DHL such that the restricted ontologies are parallelly tractable.
We further extend the results to an extension of DHL that also allows \emph{complex role inclusion axioms}.
To verify the practical usability of the above results, we analyze real-world datasets and show
that many ontologies expressed in DHL or its extension belong to
the parallelly tractable classes. Based on the results of parallel tractability,
we further implement an optimized parallel algorithm
and compare it to the state-of-the-art reasoner RDFox. The experimental results show that the
optimization used in our implementation results in a better
performance on parallelly tractable ontologies compared with RDFox.
\end{abstract}

>>>>

\address[add2]{
    Institute of Artificial Intelligence, University of Ulm
}

\begin{abstract}
  Materialization is an important reasoning service for many
  ontology-based applications, but the rapid growth of semantic data
  poses the challenge to efficiently perform materialization on
  large-scale ontologies. Parallel materialization algorithms work
  well for some ontologies, which is not always explained by the
  theoretical results. For example, for the YAGO ontology
  parallelization works well, although the reasoning problem for the
  used ontology language is not in \texttt{NC}, i.e.\ the theoretical
  complexity class for \emph{parallel tractability}. This motivates us
  to study the problem of \emph{parallel tractability} of ontology
  materialization from a theoretical perspective. We focus on the
  datalog rewritable ontology languages DL-Lite and Description Horn
  Logic (DHL) and propose algorithms, called \texttt{NC} algorithms,
  to identify classes of ontologies for which materialization is
  tractable in parallel. To verify the practical usability of the
  above results, we analyze real-world datasets and show that for many
  ontologies expressed in DHL materialization is tractable in
  parallel. The implementation of our optimized parallel algorithm
  shows performance improvements over the highly optimized
  state-of-the-art reasoner RDFox on ontologies for which
  materialization is tractable in parallel.
\end{abstract}
